FLM for growth as function of SPEI

### 2021-02-19 analysis/paper/vital rate models.Rmd
rotated crossbasis plots, added line for SPEI = 0, added season bar, made difference rasters.

`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

TODO: It would be *very cool* if the colors could be the same, but the limits on the scale bar could be different.

```{r}
s_1ha_cbplot <-
  ggplot(s_1ha_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  scale_fill_viridis_c("P(survival)", option = "magma", limits = surv_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) +
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "1ha fragments") +
  theme_bw()
ggsave(here("analysis", "figures", "surv_1ha_spei.png"), s_1ha_cbplot)
```
```{r}
s_cf_cbplot <- 
  ggplot(s_cf_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  scale_fill_viridis_c("P(survival)", option = "magma", limits = surv_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) + 
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "Continuous forest") +
  theme_bw()
ggsave(here("analysis", "figures", "surv_cf_spei.png"), s_cf_cbplot)
```

combine
```{r}
s_1ha_cbplot + s_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "surv_both_spei.png"), width = 10, height = 5)
```
Try rotating to put time on x-axis

```{r}
breaks <- seq(0,36, by = 2)
ggplot(s_1ha_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  scale_fill_viridis_c("P(survival)", option = "magma", limits = surv_lims) +
  scale_x_reverse(
    "lag (months)",
    breaks = breaks,
    expand = c(0, 0),
    labels = function(x) -x,
    sec.axis = dup_axis(name = "month",
                        labels = month(ymd("2020-2-01") - months(breaks), label = TRUE))
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  # labs(title = "1ha fragments") +
  theme_bw() +
  theme(axis.text.x.top = element_text(angle = 45, vjust = 0.2, hjust = 0)) +
  NULL
```

@@

## Crossbasis plot
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

```{r}
g_1ha_eval <-
  evaluate_smooth(g_post_1ha, smooth = "spei_history", dist = 0.1) #%>% 
  # mutate(est = est + coef(g_post_1ha)[1])
g_cf_eval <-
  evaluate_smooth(g_post_cf, smooth = "spei_history", dist = 0.1) # %>% 
  # mutate(est = est + coef(g_post_cf)[1])

@@

```{r}
g_1ha_cbplot <- 
  ggplot(g_1ha_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.1) +
  scale_fill_viridis_c("∆log(size) (?)", option = "magma", limits = grow_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) +
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "1ha fragments") +
  theme_bw()
ggsave(here("analysis", "figures", "growth_1ha_spei.png"), g_1ha_cbplot)
```

```{r}
g_cf_cbplot <-
  ggplot(g_cf_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.1) +
  scale_fill_viridis_c("∆log(size) (?)", option = "magma", limits = grow_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) +
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "Continuous forest") +
  theme_bw()
ggsave(here("analysis", "figures", "growth_cf_spei.png"), g_cf_cbplot)
```

combine

```{r}
g_1ha_cbplot + g_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "growth_both_spei.png"), width = 10, height = 5)
```

@@

```{r}
bind_rows("1-ha" = f_1ha_eval1, "CF" = f_cf_eval1, .id = "habitat") %>% 
  ggplot(aes(x = log_size_prev)) +
  geom_line(aes(y = est, color = habitat)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, fill = habitat), alpha = 0.25) +
  # stat_summary_bin(geom = "point", fun = mean, aes(y = flwr), data = model.frame(f_post_cf)) +
  geom_rug(data = bind_rows(
    "1-ha" = model.frame(f_post_1ha),
    "CF" = model.frame(f_post_cf),
    .id = "habitat")) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = "P(flowering)")
ggsave(here("analysis", "figures", "flwr_both_size.png"))
```

`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

```{r}
f_1ha_cbplot <- 
  ggplot(f_1ha_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  scale_fill_viridis_c("P(flowering)", option = "magma", limits = flwr_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) +
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "1-ha fragments") +
  theme_bw()
ggsave(here("analysis", "figures", "flwr_1ha_spei.png"), f_1ha_cbplot)
```
```{r}
f_cf_cbplot <- 
  ggplot(f_cf_eval, aes(x = spei_history, y = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  scale_fill_viridis_c("P(flowering)", option = "magma", limits = flwr_lims) +
  scale_y_continuous(
    "lag (months)",
    breaks = 0:36, 
    expand = c(0,0),
    sec.axis = dup_axis(
      name = "month",
      labels = month(ymd("2020-2-01") - months(0:36), label = TRUE))
  ) + 
  scale_x_continuous(TeX("SPEI_3_"), expand = c(0,0)) +
  labs(title = "Continuous forest") +
  theme_bw()
ggsave(here("analysis", "figures", "flwr_cf_spei.png"), f_cf_cbplot)
```

combine
```{r}
f_1ha_cbplot + f_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "flwr_both_spei.png"), width = 10, height = 5)
```

>>>>

### Setup for all crossbasis plots
Ok, I want to add a lil strip on the bottom that shows dry and wet seasons.  Wet = November-may, dry = june-october.

```{r}
wet_xmaxs = c(3, 15, 27, 36)
wet_xmins = c(0, 8, 20, 32)
dry_xmaxs = c(8, 20, 32)
dry_xmins = c(3, 15, 27)
df <- tibble(x = 0:36)
season_bar <- 
  ggplot(df, aes(x = x, y = 1)) +
  annotate(geom = "rect", xmin = wet_xmins, xmax = wet_xmaxs, ymin = 0.975, ymax = 1.025, fill = "black") +
  annotate(geom = "rect", xmin = dry_xmins, xmax = dry_xmaxs, ymin = 0.975, ymax = 1.025, fill = "white") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  theme_void()
```

```{r}
breaks <- seq(0,36, by = 2)
```

### Get plot data
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

### 1ha

```{r}
s_1ha_rotated <-
  ggplot(s_1ha_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(survival)", option = "viridis", limits = surv_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  ) +
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "1ha fragments") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```

### CF

```{r}
s_cf_rotated <-
  ggplot(s_cf_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(survival)", option = "viridis", limits = surv_lims) +
  scale_x_continuous(
    "lag (months)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0))) +
  labs(title = "Continuous Forest") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL

```

### Combine

```{r}
(s_cf_rotated + theme(axis.title.x = element_blank())) / 
  (s_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "surv_both_spei.png"), width = 7, height = 5)
```

### Difference
Try plotting a raster of the difference between CF and 1-ha

```{r}
s_diff_rotated <- 
  bind_cols(
  s_cf_eval %>% rename_with(.fn = ~glue("cf_{.}")),
  s_1ha_eval %>% rename_with(.fn = ~glue("frag_{.}"))
  ) %>%
  mutate(diff = cf_est - frag_est) %>% 
  ggplot(aes(x = cf_L, y = cf_spei_history, fill = diff, z = diff)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.002) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("∆Survival (CF – 1ha)", option = "viridis") +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0))) +
  labs(title = "CF – 1-ha") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
ggsave(here("analysis", "figures", "surv_diff_spei.png"),plot = s_diff_rotated, width = 7, height = 2.5)
```

@@

### Get plot data
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

```{r}
g_1ha_eval <-
  evaluate_smooth(g_post_1ha, smooth = "spei_history", dist = 0.1) %>% 
  mutate(est = est + coef(g_post_1ha)[1])
g_cf_eval <-
  evaluate_smooth(g_post_cf, smooth = "spei_history", dist = 0.1)  %>% 
  mutate(est = est + coef(g_post_cf)[1])

@@

### 1ha

```{r}
g_1ha_rotated <-
  ggplot(g_1ha_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c(TeX("$log(size_{t+1})$"), option = "viridis", limits = grow_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "1ha fragments") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```

### CF

```{r}
g_cf_rotated <-
  ggplot(g_cf_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c(TeX("$log(size_{t+1})$"), option = "viridis", limits = grow_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "Continuous forest") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```

### Combine
combine
```{r}
(g_cf_rotated + theme(axis.title.x = element_blank())) / 
  (g_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "growth_both_spei.png"), width = 7, height = 5)
```
### Difference

```{r}
g_diff_rotated <- 
  bind_cols(
  g_cf_eval %>% rename_with(.fn = ~glue("cf_{.}")),
  g_1ha_eval %>% rename_with(.fn = ~glue("frag_{.}"))
  ) %>%
  mutate(diff = cf_est - frag_est) %>% 
  ggplot(aes(x = cf_L, y = cf_spei_history, fill = diff, z = diff)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.05) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c(TeX("$\\Delta log(size_{t+1})$ (CF – 1ha)"), option = "viridis") +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0))) +
  labs(title = "CF – 1-ha") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
ggsave(here("analysis", "figures", "growth_diff_spei.png"),plot = g_diff_rotated, width = 7, height = 2.5)
```

@@

```{r}
bind_rows("1-ha" = f_1ha_eval1, "CF" = f_cf_eval1, .id = "habitat") %>% 
  ggplot(aes(x = log_size_prev)) +
  geom_line(aes(y = est, color = habitat)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci, fill = habitat), alpha = 0.25) +
  # stat_summary_bin(geom = "point", fun = mean, aes(y = flwr), data = model.frame(f_post_cf)) +
  geom_rug(data = bind_rows(
    "1-ha" = model.frame(f_post_1ha),
    "CF" = model.frame(f_post_cf),
    .id = "habitat")) +
  theme_bw() +
  labs(x = TeX("$log(size_t)$"), y = "P(flowering)")
ggsave(here("analysis", "figures", "flwr_both_size.png"))
```
```{r paged.print=FALSE}
demog_full %>% filter(ht < 1)
```

## Crossbasis plot
### Get plot data
`dist` isn't working yet in `smooth_estimates`, but that's ok because I don't need CI for these plots (yet).

@@

### 1ha

```{r}
f_1ha_rotated <-
  ggplot(f_1ha_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(flowering)", option = "viridis", limits = flwr_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "1ha fragments") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```

### CF
```{r}
f_cf_rotated <-
  ggplot(f_cf_eval, aes(y = spei_history, x = L, fill = est, z = est)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("P(flowering)", option = "viridis", limits = flwr_lims) +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0)))+ #leave room for season bar at bottom
  labs(title = "Continuous forest") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
```
### Combine
combine
`
```{r}
  (f_cf_rotated + theme(axis.title.x = element_blank())) / 
  (f_1ha_rotated) + 
  plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = "a", tag_suffix = ")") & 
  labs(title = NULL) #remove plot titles
ggsave(here("analysis", "figures", "flwr_both_spei.png"), width = 7, height = 5)
```
### Difference

```{r}
f_diff_rotated <- 
  bind_cols(
  f_cf_eval %>% rename_with(.fn = ~glue("cf_{.}")),
  f_1ha_eval %>% rename_with(.fn = ~glue("frag_{.}"))
  ) %>%
  mutate(diff = cf_est - frag_est) %>% 
  ggplot(aes(x = cf_L, y = cf_spei_history, fill = diff, z = diff)) +
  geom_raster() +
  geom_contour(color = "black", binwidth = 0.001) +
  geom_hline(aes(yintercept = 0), color = "grey", linetype = 2) +
  scale_fill_viridis_c("∆Flowering (CF – 1ha)", option = "viridis") +
  scale_x_continuous(
    "lag (months before census)",
    breaks = breaks,
    expand = c(0, 0)
  )+
  scale_y_continuous(TeX("SPEI_3_"), expand = expansion(mult = c(0.025, 0))) +
  labs(title = "CF – 1-ha") +
  theme_bw() +
  annotation_custom(grob = ggplotGrob(season_bar), ymin = -Inf, ymax = min(s_cf_eval$spei_history), xmin = -Inf, xmax = Inf) +
  NULL
ggsave(here("analysis", "figures", "flwr_diff_spei.png"),plot = s_diff_rotated, width = 7, height = 2.5)
```
